Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2021-06-07 main.tex
Readthrough with little fixes

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private and navigator estimates cannot be learned by sensors. We first define a novel notion for private linear combination aggregation, satisfying these cryptographic requirements, before using it to derive a filter based on the Extended Kalman Filter, with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition capturing communication assumptions and leakages is given, crucial for cryptographic security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

We motivate this scenario with the example of vehicle localisation in the presence of privately owned measurement stations. While the intention of measurement stations is the accurate navigation of passing vehicles, it may be reasonable to desire identifying location or hardware details to remain unknown to the other present stations and to the navigator. Similarly, a navigator may not wish to disclose their most accurate location estimates to untrusted third-party measurers.

In section \ref{sec:problem_statement}, we introduce both the cryptographic and estimation problems considered in this work, before giving some relevant preliminaries in section \ref{sec:preliminaries}. Section \ref{sec:lcao_scheme} proposes a cryptographic scheme meeting our desired security properties and section \ref{sec:priv_localization} the solution to the estimation problem making use of this scheme. Simulation results are discussed in section \ref{sec:sim_and_results} and concluding remarks are made in section \ref{sec:conclusion}.

@@

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\hat{\vec{x}}_{k|l}$ will denote an estimate of $\vec{x}$ at timestep $k$ given measurements from timesteps up to and including $l$. When discussing cryptography, $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$ will denote encryption and decryption with key $k$, respectively, with $k$ ommited when inferrable from context. $\mathbb{Z}_n$ is used for the set of integers m  odulo $n$ and $\mathbb{Z}^*_n$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ the binary concatination operator and the term \textit{negligible function} refers to the cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] Aggregation is computed as $l^{(t)} = \prod^n_{i=1}l^{(t)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(t)}_{i,j}\omega^{(t)}_j =\\
            &\frac{L((l^{(t)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        l^{(t)} &= \prod^n_{i=1}H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}a_{i,j}\omega_j}\rho_i^N \pmod{N^2} \\
        \begin{split}
            &= H(t)^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\rho_i^N \pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\rho'^N \pmod{N^2}\,,
    \end{align*}
    for some values $\rho_i,\rho' \in \mathbb{Z}_N,\,1\leq i \leq n$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(t)}_j$ and values $a^{(t)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed $\pmod{N}$.

@@

\subsection{Leakage} \label{subsec:leakage}
With the privacy-preserving EIF defined in the previous section, we can now interpret the aggregation leakage of an LCAO scheme in the context of range sensor localisation. The leakage function from the $\mathsf{AggDec}$ algorithm corresponds to the information vector and matrix sums, $\sum_{i=1}^n\vec{i}'_{k,i}$ and $\sum_{i=1}^n\mat{I}'_{k,i}$, respectively. However, recalling that a compromised navigator can also learn the individual sums weighted by the same weight, which in this case means the leakage of $\{\sum_{i=1}^n2r^{-1}_{k,i},\ \sum_{i=1}^n-r^{-1}_{k,i}s_{x,i},\ \sum_{i=1}^n-2r^{-1}_{k,i}s_{x,i},\ \dots\}$. From these leaked sums, we can see that private sensor information, $z'_{k,i}$, $r'_{k,i}$ and $\vec{s}_i$, is present only in their complete sums
\begin{equation}
    \sum_{i=1}^nz'_{k,i}\,,\ \sum_{i=1}^nr'_{k,i}\,,\ \sum_{i=1}^ns_{x,i}\ \textrm{and}\ \sum_{i=1}^ns_{y,i}\,,\label{eqn:localisation_leakage}
\end{equation}
which in practice correspond to their averages. We can therefore say, in the context of our proposed localisation method, LCAO leakage corresponds to the averages of sensor private information, while individual sensor information remains private.

@@

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. The layouts, of varying sensor distances to the navigator track, are shown in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Diffrent simulation layouts with varying average distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
To demonstrate the accuracy of the method, we have compared the root mean square error (RMSE) of the privacy-preserving filter to the standard EIF using unmodified measurements, that is algebraically equivalent to the EKF, typically used in industry for linearising non-linear state estimation. Measurement and estimation in each layout from figure \ref{fig:sim_layouts} consists of $50$ filter iterations and was run $100$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{MSE of the privacy-preserving EIF and standard EIF for different layout simulations.}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see the similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and navigator have little impact on the differences in performance. We can attribute the similarity in MSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.

>>>>

Our contribution in this work presents a range-only localisation method meeting formal cryptographic requirements that ensure sensors keep their measurements and locations private while the navigator keeps its estimates private. We first define a novel notion for private linear combination aggregation, satisfying these cryptographic requirements, before using it to derive a filter based on the Extended Kalman Filter, with no hierarchical sensor layout assumptions. The linear combination aggregation scheme we put forward is in principle similar to the pWSAh scheme in \cite{alexandruPrivateWeightedSum2020}, however, a formal definition capturing communication assumptions and leakages is given, crucial for cryptographic security. To the best of the authors' knowledge, no existing method for Bayesian state estimation using range-only sensors and meeting the desired privacy requirements exists.

We motivate this scenario with the example of vehicle localisation in the presence of privately owned measurement stations. While the intention of measurement stations is the accurate navigation of passing vehicles, it may be reasonable to desire identifying location or hardware details to remain unknown to the other present stations and the navigator. Similarly, a navigator may not wish to disclose their most accurate location estimates to untrusted third-party measurers.

In section \ref{sec:problem_statement}, we introduce both the cryptographic and estimation problems considered in this work, before giving some relevant preliminaries in section \ref{sec:preliminaries}. Section \ref{sec:lcao_scheme} proposes a cryptographic scheme meeting our desired security properties and section \ref{sec:priv_localization} gives a solution to the estimation problem making use of this scheme. Simulation results are discussed in section \ref{sec:sim_and_results} and concluding remarks are made in section \ref{sec:conclusion}.

@@

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\hat{\vec{x}}_{k|l}$ will denote an estimate of $\vec{x}$ at timestep $k$ given measurements from timesteps up to and including $l$. When discussing cryptography, $\mathcal{E}_{k}(\cdot)$ and $\mathcal{D}_{k}(\cdot)$ will denote encryption and decryption with key $k$, respectively, with $k$ ommited when inferrable from context. $\mathbb{Z}_n$ is used for the set of integers modulo $n$ and $\mathbb{Z}^*_n$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ the binary concatination operator and the term \textit{negligible function} refers to its cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\shrtdots,l^{(t)}_n)$] Aggregation is computed as $l^{(t)} = \prod^n_{i=1}l^{(t)}_i \pmod{N^2}$, removing the aggregation noise terms, and is followed by Paillier scheme decryption
    \begin{equation}
        \begin{split}
            \sum^{n}_{i=1}\sum^{m}_{j=1}&a^{(t)}_{i,j}\omega^{(t)}_j =\\
            &\frac{L((l^{(t)})^\lambda\pmod{N^2})}{L((N+1)^\lambda\pmod{N^2})} \pmod{N}\,,
        \end{split} \label{eqn:our_scheme_decrypt}
    \end{equation}
    with $\lambda = \mathsf{lcm}(p-1, q-1)$ and $L(u) = \frac{u-1}{N}$. The correctness of the aggregation can be seen from
    \begin{align*}
        l^{(t)} &= \prod^n_{i=1}H(t)^{sk_i}(N+1)^{\sum^{m}_{j=1}a_{i,j}\omega_j}\rho_i^N \pmod{N^2} \\
        \begin{split}
            &= H(t)^{\sum^n_{i=1}sk_i}\cdot \\
            &\qquad\qquad \prod^n_{i=1}(N+1)^{\sum^{m}_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\rho_i^N \pmod{N^2}
        \end{split}\\
        &= (N+1)^{\sum^n_{i=1}\sum^{m}_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j}\rho'^N \pmod{N^2}\,,
    \end{align*}
    for some values $\rho_i,\rho' \in \mathbb{Z}_N,\,1\leq i \leq n$.
\end{LaTeXdescription}
Additionally, we note that in the above construction, all weights $\omega^{(t)}_j$ and values $a^{(t)}_{i,j}$ are integers and the resulting linear combinations and aggregation are computed modulo $N$.

@@

\subsection{Leakage} \label{subsec:leakage}
With the privacy-preserving EIF defined in the previous section, we can now interpret the aggregation leakage of an LCAO scheme in the context of range sensor localisation. The leakage function from the $\mathsf{AggDec}$ algorithm corresponds to the information vector and matrix sums, $\sum_{i=1}^n\vec{i}'_{k,i}$ and $\sum_{i=1}^n\mat{I}'_{k,i}$, respectively. However, recalling that a compromised navigator can also learn the individual sums weighted by the same weight, $\{\sum_{i=1}^n2r^{-1}_{k,i},\ \sum_{i=1}^n-r^{-1}_{k,i}s_{x,i},\ \sum_{i=1}^n-2r^{-1}_{k,i}s_{x,i},\ \dots\}$ can be leaked as well. From these leaked sums, we can see that private sensor information, $z'_{k,i}$, $r'_{k,i}$ and $\vec{s}_i$, is present only in their complete sums
\begin{equation}
    \sum_{i=1}^nz'_{k,i}\,,\ \sum_{i=1}^nr'_{k,i}\,,\ \sum_{i=1}^ns_{x,i}\ \textrm{and}\ \sum_{i=1}^ns_{y,i}\,,\label{eqn:localisation_leakage}
\end{equation}
which in practice correspond to their averages. We can therefore say, in the context of our proposed localisation method, LCAO leakage corresponds to the averages of sensor private information, while individual sensor information remains private.

@@

We have considered multiple sensor layouts, each with four sensors, to capture the dependence of estimated modified measurement variances $r'_{k,i}$ on the original measurements $z_{k,i}$. These layouts of varying sensor distances to the navigator track are shown in figure \ref{fig:sim_layouts}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layouts.pdf}
    \caption{Diffrent simulation layouts with varying average distances between navigator and sensors.}
    \label{fig:sim_layouts}
\end{figure}
To demonstrate the accuracy of the method, we have compared the root mean square error (RMSE) of the privacy-preserving filter to the standard EIF using unmodified measurements, that is algebraically equivalent to the EKF, typically used in industry for linearising non-linear state estimation. Measurement and estimation in each layout from figure \ref{fig:sim_layouts} consists of $50$ filter iterations and was run $100$ times. Unmodified measurement variances were taken as $r_{k,i}=5$ for all $k>0$ and a large fractional precision factor, $\phi=2^{32}$, was chosen. The results can be seen in figure \ref{fig:sim_layout_errors}.
\begin{figure}[htbp]
    \centering
    \includegraphics{images/layout_errors.pdf}
    \caption{MSE of the privacy-preserving EIF and standard EIF for different layout simulations.}
    \label{fig:sim_layout_errors}
\end{figure}
From these results, we can see the similarity in filter performance between the privacy-preserving method and that of the traditional EIF. We can also see that the varying average distances between sensors and navigator have little impact on the differences in performance. We can attribute the similarity in MSE to the conservativeness of estimated modified measurement variances $r'_{k,i}$, resulting in few additional filter divergences, and to the high fractional precision factor, keeping computations consistent with the floating-point arithmetic of the EIF.
